SVG: The Paper (2nd part of supplementary material)

### 2022-06-07 Analysis/index2.Rmd
testing and training

---
title: 'SVG: The Paper'
author: "Alasdair Clarke, Amelia Hunt &  Anna Hughes"
date: "13/04/2022"
output:
  tufte::tufte_html:
    number_sections: true
    fig_height: 4
    bibliography: literature.bib
    link-citations: yes  
---

@@

We will use the visual foraging data from @clarke2022 to assess the model's trial level performance. As the current version of our model only takes the $p_A$ bias into account for the initial target selection, we will ignore these data points for now.

```{r, echo = TRUE}
# read in data
d <- read_csv("data/clarke_2020_qjep.csv", show_col_types = FALSE) %>%
  mutate(condition = as_factor(condition),
         condition = fct_recode(condition, feature = "1", conjunction = "2"),
         targ_type = as_factor(targ_type))
```

@@

```{r echo = TRUE}
# item weights pre-computed - see scripts/extract_item_weights_using_model.R
a <- readRDS("scratch/qjep_model_weights.rda") %>%
  filter(found != 1) # remove initial selections

@@

ggsave("../Figures/qjep_preds.pdf", 
       plt_b + plt1 + plt2 + 
         plot_layout(guides = "collect") & theme(legend.position = "bottom"),
       width = 10, height = 4)

rm(plt1, plt2, plt_a)
```

@@

```{r fig.height = 6, fig.cap = "Accuracy of our model varies with the strength of an individual's model parameters. We can see two clear outlier participants (marked with an X). "}
# computed in xxxx.R
fit <- readRDS("scratch/qjep_model_fit.rda") %>%
  full_join(a_acc, by = c("observer", "condition")) %>%
  mutate(is_outlier = if_else(observer %in% c(35, 45), TRUE, FALSE))

@@

ggsave("../Figures/qjep_indiv_diff.pdf", 
       plt_a + plt_b + 
         plot_layout(guides = "collect") & theme(legend.position = "right"), width = 8, height = 4)
```

As the two outlier participants appear to be doing something different, not captured especially well by our model, we will remove them from the rest of this analysis.

After removing them, we can see that individual differences in proximity tuning account for nearly all of the differences in predictability from person to person, and also between the *feature* and *conjunction* conditions.

```{r, echo = TRUE}
dm <- filter(fit, is_outlier == 0)

@@

Both $p_A$ and $b_M$ also have an effect on our model's accuracy, but note there is only a small increases in $R^2$ from just using $b_P$.

## Are more predictable people faster at the task?

```{r}
drt <- filter(d, found == 40) %>% group_by(observer, condition) %>%
  summarise(median_rt = median(RT), .groups = "drop") %>% 
  full_join(dm, by = c("observer", "condition"))

ggplot(drt, aes(accuracy, median_rt, colour = condition)) +
         geom_point(alpha = 0.5) + 
  geom_smooth(method = "lm")

summary(lm(median_rt ~ accuracy*condition, data = drt))

```{r}
rm(dm)
```

## Examples of Trials

Let's score every trial by how well the model captures behaviour, and then plot some different examples.

```{r, echo = TRUE}
a %>% filter(!(observer %in% c(35, 45))) %>%
  group_by(observer, condition, trial) %>% 
  summarise(prop_max = mean(selected_max), .groups = "drop") %>% 
  arrange(desc(prop_max)) %>%
  ungroup() -> a_trl
```

```{r fig.height=3.5, fig.cap = "Histogram of how predictiable individual trials are, defined as how often the model assigned the largest weight to the target that was then selected."}
ggplot(a_trl, aes(prop_max)) + 
  geom_histogram(fill = "#d04e00", breaks = seq(0, 1, 0.025), alpha = 0.5) +
  scale_x_continuous("trial predictability")
```

Let's plot some examples oh highly predictable (top 5%), typical (middle 10%) and unpredictable (bottom 5%) trials.

```{r, echo = TRUE}
qs <- quantile(a_trl$prop_max, c(0.05, 0.40, 0.60, 0.95))

trls_hard <- filter(a_trl, prop_max < qs[1]) %>% sample_n(3)
trls_mid  <- filter(a_trl, prop_max > qs[2], prop_max < qs[3]) %>% sample_n(3)
trls_easy <- filter(a_trl, prop_max > qs[4]) %>% sample_n(3)
```

We can see that for these highly predictable trials, we make very few mistakes, and those that we do make appear to be related to some sort of shortest-path optimization on the part of the participant.

```{r fig.cap = "A random selection of *good* trials. When participants diverge from the model's prediction, it appears to be due to some form of path-length optimisation."}

source("functions/plot_trial.R")
plts1 <- map(1:3, plot_trial, trls_easy, a, d)

plts1[[1]] + plts1[[2]] +
  plot_layout(nrow = 1, widths = c(1,1)) &
   theme(plot.margin = unit(c(.2,.2,.2,.2), "cm")) &
   theme(plot.background = element_rect(color  = "#fffff8", size = 2,linetype = 'dotted', fill ="#fffff8")) &
   theme(text=element_text(size=14,  family="serif"))
```

Here are some typical trials. Possible we have the path-length minimization issue again.

```{r fig.cap = "A random selection of *typical* trials. Along with path-minimising behaviour, we can also see some disagreement around switching item type."}
plts2 <- map(1:3, plot_trial, trls_mid, a, d)

plts2[[1]] + plts2[[2]] +
  plot_layout(nrow = 1, widths = c(1,1)) &
   theme(plot.margin = unit(c(.2,.2,.2,.2), "cm")) &
   theme(plot.background = element_rect(color  = "#fffff8", size = 2,linetype = 'dotted', fill ="#fffff8")) &
   theme(text=element_text(size=14,  family="serif"))
```

Now let's look at less predictable (bottom 5%-tile) trials. Even here, the model does a relatively good job.

```{r fig.cap = "A random selection of *hard* trials. It is harder to make out much of a pattern here, but we can see that the human participants jump around more."}
plts3 <- map(1:3, plot_trial, trls_hard, a, d)

plts3[[1]] + plts3[[2]] +
  plot_layout(nrow = 1, widths = c(1,1)) &
   theme(plot.margin = unit(c(.2,.2,.2,.2), "cm")) &
   theme(plot.background = element_rect(color  = "#fffff8", size = 2,linetype = 'dotted', fill ="#fffff8")) &
   theme(text=element_text(size=14,  family="serif"))
```

```{r}
ggsave("../Figures/qjep_ex_paths.pdf",
      ( plts1[[1]] + plts1[[2]]) /( plts3[[1]] + plts3[[3]]), 
       width = 8, height = 8)

rm(trls_easy, trls_mid, trls_hard, plts, qs, plts1, plts2, plts3)
```

# Path Length Analysis

It looks like some of the target selections that our model fails to capture are when participants select a further away target in order to minimise overall path length. Here we will compare our model's predicted path length against the empirical paths.

- We will provide the model/simulation with the initial target selection.

```{r, fig.height=3.5, fig.cap = "(*left*) Histogram comparing the empirical path lengths to simulations from our fitted model. Difference indicates the human path length - the simulated path length. (*right*) Histogram comparing the median path lengths per participant.", message=FALSE}
path_lengths <- readRDS("scratch/path_lengths.rda") %>%
  rename(observer = "obs",
         condition = "cond",
         human = "human_path_length",
         model = "model_path_length") %>% 
  mutate(difference  = human - model) %>%
  pivot_longer(c(human, model, difference), names_to = "type", values_to = "path_length")

ggplot(path_lengths, aes(path_length, fill = type)) + 
  geom_histogram(bins = 25, alpha = 0.5, position = "identity") +
  geom_vline(xintercept = 0, linetype = 2) + 
  theme(legend.position = "bottom")  +
  scale_x_continuous("path length")  +
  scale_x_continuous("median path length") +
  scale_fill_manual(values = col3) -> plt1

path_lengths %>% group_by(observer, type) %>%
  summarise(path_length = median(path_length), .groups = "drop") %>%
  ggplot(aes(path_length, fill = type)) + 
  geom_histogram(bins = 25, alpha = 0.5, position = "identity") +
  geom_vline(xintercept = 0, linetype = 2) + 
  theme(legend.position = "bottom") +
  scale_x_continuous("median path length") +
  scale_fill_manual(values = col3) -> plt2

plt1 + plt2 + plot_layout(guides = "collect")  &
  theme(legend.position = 'bottom', legend.title = element_blank())
```

We can see that overall, the model is not making substantially longer paths through the items than our human participants. However, there is some variability in that some participants routinely select paths shorter than our model would, and others do not.

Perhaps this is related to predictability (i.e., how well the model can predict a participant's trial level behaviour)?

```{r fig.height=3, fig.cap = "(*left*) A histogram showing the proportion of trials each participant managed a shorter path through the targets than the model's prediction. (*right*) Scatter plot showing the relationship between predictability (accuracy) and whether the model over or under estimates path length. Xs mark the outlier participants discussed above."}
path_lengths %>% filter(type == "difference") %>%
  mutate(shorter = path_length < 0) %>%
  group_by(observer, condition) %>%
  summarise(prop_shorter = mean(shorter), .groups = "drop") -> path_lengths_obs

path_lengths_obs %>%
  ggplot(aes(prop_shorter, fill = condition)) + 
  geom_histogram(alpha = 0.5, breaks = seq(0, 1, 0.2), position = "identity") +
  scale_x_continuous("prop. shorter than model", breaks = seq(0, 1, 0.2)) -> plt1

path_lengths <- full_join(path_lengths_obs, fit,
                          by = c("observer", "condition"))

ggplot(path_lengths, aes(accuracy, prop_shorter, colour = condition)) + 
  geom_point(aes(shape = is_outlier)) +
  geom_smooth(data = filter(path_lengths, is_outlier == 0), 
              method = "lm", formula = y ~ x, se = FALSE) + 
  scale_shape_manual(values = c(19, 4), guide = "none") + 
  scale_y_continuous("prop. shorter than model", breaks = seq(0, 1, 0.2)) +
  coord_cartesian(ylim = c(0, 1)) -> plt2

plt1 + plt2 + plot_layout(guides = "collect")  &
  theme(legend.position = 'bottom')
```

```{r}
rm(path_lengths, plt1, plt2)
```

@@

```{r echo = TRUE}
# Reimport data again (as we removed the initial target selection earlier)
d <- read_csv("data/clarke_2020_qjep.csv", show_col_types = FALSE) %>%
  filter(found == 1) %>%
  mutate(condition = as_factor(condition))
```

@@

plt_all + plt_med 
ggsave("../Figures/init_sel_hex_plot.pdf", width = 8, height = 4)

@@

## Multi-level Model

We can model this as multi-level beta distribution? Yes, although it seems like people are in one of two clusters.

```{r fig.cap = "The mean posterior plot for the truncated normal distriubtions for (*left*) $x$ and (*right*) $y$. We can see that a two component mixture model will capture most of the variance in intial target selection."}

m <- readRDS("scratch/init_sel_model.rds")

gather_draws(m, u[dim, observer]) %>% 
  ungroup() %>%
  select(-.variable) %>%
  rename(obs_b = ".value") %>%
  left_join(gather_draws(m, b[dim]), 
            by = c("dim", ".chain", ".iteration", ".draw")) %>%
  mutate(obs_b = exp(.value + obs_b),
         .variable = if_else(dim %in% c(1,2), "a", "b"),
         dim = if_else(dim %in% c(1,3), "x", "y")) %>%
  select(observer, dim, .iteration, obs_b, param = ".variable") -> a

a %>% group_by(observer, param, dim) %>%
  summarise(value = mean(obs_b) , .groups = "drop") %>%
  pivot_wider(names_from = param, values_from = value) -> am

compute_beta_dist <- function(observer, a, b) {
  x <- seq(0.01,0.99,0.01)
  return(tibble(observer = observer, 
                x=  x,
                z = dbeta(x, a, b)))

}

create_post_plot <- function(dm) {

   pmap_df(filter(am,  dim == dm) %>% select(observer, a, b), 
                      compute_beta_dist) %>%
    ggplot(aes(x = x, y = z, group = observer)) + 
    geom_path(alpha = 0.25) +
    scale_x_continuous(dm) + 
    scale_y_continuous("likelihood") -> plt

  return(plt)
}

create_post_plot("x") + create_post_plot("y") 
ggsave("../Figures/init_sel_mdl.pdf", width = 10, height = 3.5)
```

## Mixture Model

It looks like we can simplify the model somewhat by using a two component mixture model instead!

```{r, fig.height = 6, fig.cap = "(*top*) Beta distributions for $x$ and $y$ coordinated"}

m <- readRDS("scratch/init_sel_model2.rds")

sample_beta <- function(c, .draw, dim, a, b) {
  x <- seq(0.01, 0.99, 0.01)
  return(tibble(c = c, 
                draw = .draw,
                dim = dim,
                x = x,
                z = dbeta(x, shape1 = a, shape2 = b)))
}

gather_draws(m, a_x[c], b_x[c], a_y[c], b_y[c], ndraws = 100) %>%
  separate(.variable, into = c("param", "dim")) %>%
  select(-.chain, -.iteration) %>%
  pivot_wider(names_from = param, values_from = ".value") %>%
  pmap_df(sample_beta) %>%
  mutate(c = as_factor(c)) -> plt_dat

plt_dat %>% filter(dim == "x") %>%
  ggplot(aes(x, z, group = interaction(c, draw), colour = c)) +
  geom_path(alpha = 0.1) + 
  facet_wrap(~dim) + 
  theme(legend.position = "none") -> plt_x

plt_dat %>% filter(dim == "y") %>%
  ggplot(aes(x, z, group = interaction(c, draw), colour = c)) +
  geom_path(alpha = 0.1) + 
  facet_wrap(~dim) + 
  theme(legend.position = "none") +
  scale_x_continuous("y") -> plt_y

gather_draws(m, lambda[observer]) %>%
  ggplot(aes(.value, group = observer)) + 
  geom_density(alpha = 0.25, fill = "grey") -> plt_lambdas

(plt_x + plt_y) / plt_lambdas

# Item Density

>>>>

---
title: 'SVG: The Paper (2nd part of supplementary material)'
author: "Alasdair Clarke, Amelia Hunt &  Anna Hughes"
date: "13/04/2022"
output:
  tufte::tufte_html:
    number_sections: true
    fig_height: 4
    bibliography: literature.bib
    link-citations: yes  
---

@@

This is the replication with Kristjansson (2014) data.

```{r, echo = TRUE}
# read in data
d <- read_csv("data/human_data_2014_prox.csv", show_col_types = FALSE) %>%
  mutate(condition = as_factor(condition),
         condition = fct_recode(condition, feature = "1", conjunction = "2"),
         targ_type = as_factor(targ_type))
```

@@

```{r echo = TRUE}
# item weights pre-computed - see scripts/extract_item_weights_using_model.R
a <- readRDS("scratch/kristjansson_model_weights.rda") %>%
  filter(found != 1) # remove initial selections

@@

#ggsave("../Figures/qjep_preds.pdf", 
#       plt_b + plt1 + plt2 + 
#         plot_layout(guides = "collect") & theme(legend.position = "bottom"),
#       width = 10, height = 4)

@@

```{r fig.height = 6, fig.cap = "Accuracy of our model varies with the strength of an individual's model parameters. We can see two clear outlier participants (marked with an X). "}
# computed in xxxx.R
fit <- readRDS("scratch/kristjansson_model_fit.rda") %>%
  full_join(a_acc, by = c("observer", "condition")) %>%
  mutate(is_outlier = if_else(observer %in% c(35, 45), TRUE, FALSE))

@@

#ggsave("../Figures/qjep_indiv_diff.pdf", 
#       plt_a + plt_b + 
#         plot_layout(guides = "collect") & theme(legend.position = "right"), width = 8, height = 4)
```

We can see that individual differences in proximity tuning account for nearly all of the differences in predictability from person to person, and also between the *feature* and *conjunction* conditions.

```{r, echo = TRUE}
dm <- filter(fit, is_outlier == 0) # have not currently defined anything as an outlier

@@

@@

```{r echo = TRUE}
# Reimport data again (as we removed the initial target selection earlier)
d <- read_csv("data/human_data_2014_prox.csv", show_col_types = FALSE) %>%
  filter(found == 1) %>%
  mutate(condition = as_factor(condition))
```

@@

plt_all + plt_med 
#ggsave("../Figures/init_sel_hex_plot.pdf", width = 8, height = 4)

@@
